---
title: "Assortative Matching Using Discrete Copula and Yule's Rho"
author: "[Thomas Coleman](http://www.hilerun.org/econ)"
output: html_notebook
---

# Assortative Matching Results Using Discrete Copula, Yule's Rank Correlation, Hellinger Dependence, and Related Metrics

#### See "Notes on Dependence, Correlation, and Discrete Copulas, with Application to Exchange Mobility and Assortative Matching" working paper at https://papers.ssrn.com/abstract=5057549 

#### This notebook is licensed under the [BSD 2-Clause License](https://opensource.org/licenses/BSD-2-Clause)


Based on 
Geenens, Gery. 2020. “Copula Modeling for Discrete Random Vectors.” Dependence Modeling 8 (1): 417–40. https://doi.org/10.1515/demo-2020-0022.
———. 2023. “Towards a Universal Representation of Statistical Dependence.” arXiv. https://doi.org/10.48550/arXiv.2302.08151.

and also 
Coleman, Thomas S. 2025. “Notes on Dependence, Correlation, and Discrete Copulas, with Application to Exchange Mobility and Assortative Matching.” SSRN. https://papers.ssrn.com/abstract=5057549. But more up-to-date at https://www.closemountain.com/papers/ExchangeMobilityNotes.pdf and assortative matching specifically at https://www.closemountain.com/papers/AssortNote.pdf 

Examining data from

- Chiappori, Pierre-Andre, Monica Costa Dias, Costas Meghir, and Hanzhe Zhang. 2024. “Changes in Marital Sorting: Theory and Evidence from the US.” https://hanzhezhang.github.io/research/2411SortingMeasures.pdf.
- Eika, Lasse, Magne Mogstad, and Basit Zafar. 2019. “Educational Assortative Mating and Household Income Inequality.” Journal of Political Economy 127 (6): 2795–2835. https://doi.org/10.1086/702018.
- Greenwood, Jeremy, Nezih Guner, Georgi Kocharkov, and Cezar Santos. 2014. “Marry Your Like: Assortative Mating and Income Inequality.” American Economic Review 104 (5): 348–53. https://doi.org/10.1257/aer.104.5.348.
- Schwartz, Christine R., and Robert D. Mare. 2005. “Trends in Educational Assortative Marriage from 1940 to 2003.” Demography (Springer Nature) 42 (4): 621–46. https://doi.org/10.1353/dem.2005.0036.



```{r}
rm(list=ls())    # starts a fresh workspace
#
library(knitr)
options(scipen=5)


# Alternative mpifp ??
# install.packages("mpifp")
require("mpifp")


# Load Copula and other various functions
knit('CopulaFunctions1.Rmd', tangle=TRUE)
source('CopulaFunctions1.R') 



```






## Chiappori, Costas Dias, Meghir, Zhang

chiappori, Pierre-Andre, Monica Costa Dias, Costas Meghir, and Hanzhe Zhang. 2024. “Changes in Marital Sorting: Theory and Evidence from the US.” https://hanzhezhang.github.io/research/2411SortingMeasures.pdf.


```{r}
 
# Table 2
pChiappori1950 <-matrix(c(.047,.039,.011,.002,0,
                           .031,.2,.071,.024,.006,
                           .009,.091,.108,.038,.01,
                           .002,.039,.058,.078,.02,
                           .001,.011,.024,.044,.037),5,5,byrow=TRUE)
# Too high by .001, reduce all proportionally:
pChiappori1950 <- pChiappori1950 / sum(pChiappori1950)
rChiappori1950 <- calcDiscreteCopula(pChiappori1950)
rHChiappori1950 <- calcDiscreteHellinger(pChiappori1950)
rAChiappori1950 <- altham_index(pChiappori1950)
colnames(pChiappori1950) <- c('HSD','HS','SC','C','PG')
rownames(pChiappori1950) <- c('HSD','HS','SC','C','PG')
rSChiappori1950 <- sort_parm(pChiappori1950)
rSChiappori1950_pbar <- sort_parm(rChiappori1950$pbar)


pChiappori1970 <-matrix(c(.045,.024,.012,.004,.001,
                           .02,.135,.078,.038,.013,
                           .006,.049,.115,.059,.024,
                           .002,.02,.048,.116,.049,
                           .001,.005,.018,.053,.066),5,5,byrow=TRUE)
# Too high by .001, reduce all proportionally:
pChiappori1970 <- pChiappori1970 / sum(pChiappori1970)
rChiappori1970 <- calcDiscreteCopula(pChiappori1970)
rHChiappori1970 <- calcDiscreteHellinger(pChiappori1970)
rAChiappori1970 <- altham_index(pChiappori1970)
colnames(pChiappori1970) <- c('HSD','HS','SC','C','PG')
rownames(pChiappori1970) <- c('HSD','HS','SC','C','PG')
rSChiappori1970 <- sort_parm(pChiappori1970)
rSChiappori1970_pbar <- sort_parm(rChiappori1970$pbar)


```

Calculate Pearson correlation weighted by original probabilities 

```{r}
tablesChiap50 <- list()
tablesChiap70 <- list()
namesChiap <- list('C&PG vs SC','PG vs C','C vs SC','C&PG vs other')

# Reduce to 'C&PG' vs 'SC'
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1950[3,3])
xx1[2,2] <- sum(pChiappori1950[4:5,4:5])
xx1[1,2] <- sum(pChiappori1950[3,4:5])
xx1[2,1] <- sum(pChiappori1950[4:5,3])
tablesChiap50[[1]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1970[3,3])
xx1[2,2] <- sum(pChiappori1970[4:5,4:5])
xx1[1,2] <- sum(pChiappori1970[3,4:5])
xx1[2,1] <- sum(pChiappori1970[4:5,3])
tablesChiap70[[1]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
# Reduce to 'PG' vs 'C'
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1950[4,4])
xx1[2,2] <- sum(pChiappori1950[5,5])
xx1[1,2] <- sum(pChiappori1950[4,5])
xx1[2,1] <- sum(pChiappori1950[5,4])
tablesChiap50[[2]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1970[4,4])
xx1[2,2] <- sum(pChiappori1970[5,5])
xx1[1,2] <- sum(pChiappori1970[4,5])
xx1[2,1] <- sum(pChiappori1970[5,4])
tablesChiap70[[2]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
# Reduce to 'C' vs 'SC'
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1950[3,3])
xx1[2,2] <- sum(pChiappori1950[4,4])
xx1[1,2] <- sum(pChiappori1950[3,4])
xx1[2,1] <- sum(pChiappori1950[4,3])
tablesChiap50[[3]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1970[3,3])
xx1[2,2] <- sum(pChiappori1970[4,4])
xx1[1,2] <- sum(pChiappori1970[3,4])
xx1[2,1] <- sum(pChiappori1970[4,3])
tablesChiap70[[3]] <- xx1 / sum(xx1)   # Normalize to sum to 1 ??
# Reduce to 'others' vs 'C&PG'
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1950[1:3,1:3])
xx1[2,2] <- sum(pChiappori1950[4:5,4:5])
xx1[1,2] <- sum(pChiappori1950[1:3,4:5])
xx1[2,1] <- sum(pChiappori1950[4:5,1:3])
tablesChiap50[[4]] <- xx1
xx1 <- matrix(0,nrow=2,ncol=2)
xx1[1,1] <- sum(pChiappori1970[1:3,1:3])
xx1[2,2] <- sum(pChiappori1970[4:5,4:5])
xx1[1,2] <- sum(pChiappori1970[1:3,4:5])
xx1[2,1] <- sum(pChiappori1970[4:5,1:3])
tablesChiap70[[4]] <- xx1



rChiappori_1950_2x2 <- lapply(tablesChiap50,calcDiscreteCopula)
rChiappori_1970_2x2 <- lapply(tablesChiap70,calcDiscreteCopula)
rHChiappori_1950_2x2 <- lapply(tablesChiap50,calcDiscreteHellinger)
rHChiappori_1970_2x2 <- lapply(tablesChiap70,calcDiscreteHellinger)
rSChiappori_1950_2x2 <- lapply(tablesChiap50,sort_parm)
rSChiappori_1970_2x2 <- lapply(tablesChiap70,sort_parm)
rSChiappori_1950_2x2_pbar <- lapply(rChiappori_1950_2x2, function(x) sort_parm(x$pbar))
rSChiappori_1970_2x2_pbar <- lapply(rChiappori_1970_2x2, function(x) sort_parm(x$pbar))



```


```{r}

Chiappori_results <- matrix(0,nrow=27,ncol=7)
colnames(Chiappori_results) = c('C&PG vs SC','PG vs C','C vs SC','C&PG vs other','all Yuled',
                                'all HDepend','Altham')
rownames(Chiappori_results) = c('OR 1950','OR 1970','diff','Yule 1950','Yule 1970','diff',
                                'Pearson 1950','Pearson 1970','diff','PearsonRank 1950','PearsonRank 1970','diff',
                                'Avg Diag 1950','Avg Diag 1970','diff','Avg Diag 1950 pbar','Avg Diag 1970 pbar','diff',
                                'Avg 1950 pbar normed','Avg 1970 pbar normed','diff',
                                'H Diag Assoc rho','H Diag Assoc rho','diff',
                                'H Diag Assoc dist','H Diag Assoc Dist','diff')

# First, populate odds ratios0
for (i in 1:length(tablesChiap50)) {
  Chiappori_results[1,i] = rChiappori_1950_2x2[[i]]$oddsratio
  Chiappori_results[2,i] = rChiappori_1970_2x2[[i]]$oddsratio
  Chiappori_results[3,i] = Chiappori_results[2,i] - Chiappori_results[1,i]
}

# Second, populate Yule rhos
for (i in 1:length(tablesChiap50)) {
  Chiappori_results[4,i] = rChiappori_1950_2x2[[i]]$YuleY
  Chiappori_results[5,i] = rChiappori_1970_2x2[[i]]$YuleY
  Chiappori_results[6,i] = Chiappori_results[5,i] - Chiappori_results[4,i]
}
Chiappori_results[4,5] = rChiappori1950$YuleY
Chiappori_results[5,5] = rChiappori1970$YuleY
Chiappori_results[6,5] = Chiappori_results[5,5] - Chiappori_results[4,5]
Chiappori_results[4,6] = rHChiappori1950$HellingDiagAssoc
Chiappori_results[5,6] = rHChiappori1970$HellingDiagAssoc
Chiappori_results[6,6] = Chiappori_results[5,6] - Chiappori_results[4,6]
Chiappori_results[4,7] = rAChiappori1950
Chiappori_results[5,7] = rAChiappori1970
Chiappori_results[6,7] = Chiappori_results[5,7] - Chiappori_results[4,7]

# Fiinally populate Pearson correlations
for (i in 1:length(tablesChiap50)) {
  Chiappori_results[7,i] = rChiappori_1950_2x2[[i]]$Pearson
  Chiappori_results[8,i] = rChiappori_1970_2x2[[i]]$Pearson
  Chiappori_results[9,i] = Chiappori_results[8,i] - Chiappori_results[7,i]
}
Chiappori_results[7,5] = rChiappori1950$Pearson
Chiappori_results[8,5] = rChiappori1970$Pearson
Chiappori_results[9,5] = Chiappori_results[8,5] - Chiappori_results[7,5]

# Fiinally populate Pearson midrank correlations and Avg Sorting Parm
for (i in 1:length(tablesChiap50)) {
  Chiappori_results[10,i] = rChiappori_1950_2x2[[i]]$PearsonRank
  Chiappori_results[11,i] = rChiappori_1970_2x2[[i]]$PearsonRank
  Chiappori_results[12,i] = Chiappori_results[11,i] - Chiappori_results[10,i]
  Chiappori_results[13,i] <- rSChiappori_1950_2x2[[i]]$AvgSortParm
  Chiappori_results[14,i] <- rSChiappori_1970_2x2[[i]]$AvgSortParm
  Chiappori_results[15,i] = Chiappori_results[14,i] - Chiappori_results[13,i]
  Chiappori_results[16,i] <- rSChiappori_1950_2x2_pbar[[i]]$AvgSortParm
  Chiappori_results[17,i] <- rSChiappori_1970_2x2_pbar[[i]]$AvgSortParm
  Chiappori_results[18,i] = Chiappori_results[17,i] - Chiappori_results[16,i]
  Chiappori_results[19,i] <- rSChiappori_1950_2x2_pbar[[i]]$AvgSortParmNorm
  Chiappori_results[20,i] <- rSChiappori_1970_2x2_pbar[[i]]$AvgSortParmNorm
  Chiappori_results[21,i] = Chiappori_results[20,i] - Chiappori_results[19,i]
  Chiappori_results[22,i] <- rHChiappori_1950_2x2[[i]]$HellingDiagAssoc
  Chiappori_results[23,i] <- rHChiappori_1970_2x2[[i]]$HellingDiagAssoc
  Chiappori_results[24,i] = Chiappori_results[23,i] - Chiappori_results[22,i]
  Chiappori_results[25,i] <- rHChiappori_1950_2x2[[i]]$HellingDiagDist
  Chiappori_results[26,i] <- rHChiappori_1970_2x2[[i]]$HellingDiagDist
  Chiappori_results[27,i] = Chiappori_results[26,i] - Chiappori_results[25,i]
}
Chiappori_results[10,5] = rChiappori1950$PearsonRank
Chiappori_results[11,5] = rChiappori1970$PearsonRank
Chiappori_results[12,5] = Chiappori_results[11,5] - Chiappori_results[10,5]
Chiappori_results[13,5] = rSChiappori1950$AvgSortParm
Chiappori_results[14,5] = rSChiappori1970$AvgSortParm
Chiappori_results[15,5] = Chiappori_results[14,5] - Chiappori_results[13,5]
Chiappori_results[16,5] = rSChiappori1950_pbar$AvgSortParm
Chiappori_results[17,5] = rSChiappori1970_pbar$AvgSortParm
Chiappori_results[18,5] = Chiappori_results[17,5] - Chiappori_results[16,5]
Chiappori_results[19,5] = rSChiappori1950_pbar$AvgSortParmNorm
Chiappori_results[20,5] = rSChiappori1970_pbar$AvgSortParmNorm
Chiappori_results[21,5] = Chiappori_results[20,5] - Chiappori_results[19,5]
Chiappori_results[22,5] = rHChiappori1950$HellingDiagAssoc
Chiappori_results[23,5] = rHChiappori1970$HellingDiagAssoc
Chiappori_results[24,5] = Chiappori_results[23,5] - Chiappori_results[22,5]
Chiappori_results[25,5] = rHChiappori1950$HellingDiagDist
Chiappori_results[26,5] = rHChiappori1970$HellingDiagDist
Chiappori_results[27,5] = Chiappori_results[26,5] - Chiappori_results[25,5]


print(1)
Chiappori_results




```






## Schwarz & Mare "Trends in Educational Assortative Marriage" 2005

Data for 1940, 1960, 1970, 1980, 1990, 2000
Categories years of schooling:
 <10 10-11 12 13-15 >=16
 
```{r}

tableSchwarzMare <- list()

xx1 <- c('<10','10-11','12','13-15','>=16')
# 1940
xx2 <- matrix(c(43.99,4.45,3.12,.78,.4,
                7.33,3.88,2.61,.69,.36,
                6.55,3.6,8.13,2.29,1.91,
                1.32,.67,1.47,1.58,1.62,
                .32,.16,.47,.54,1.75),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
#rownames(xx2) <- xx1
#colnames(xx2) <- xx1
tableSchwarzMare[[1]] <- xx2
# 1960
xx2 <- matrix(c(16.02,3.41,3.49,0.82,0.23,
                6.23,4.35,4.58,1.4,.46,
                7.97,6.14,17.12,6.1,3.77,
                .97,.86,2.67,3.57,3.93,
                .18,.18,.61,.97,3.95),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
#rownames(xx2) <- xx1
#colnames(xx2) <- xx1
tableSchwarzMare[[2]] <- xx2
# 1970
xx2 <- matrix(c(7.94,2.41,3.09,.69,.21,
                4.13,3.63,4.88,1.37,.38,
                5.77,5.47,21.86,8.27,4.44,
                .71,.85,3.38,5.27,5.48,
                .2,.18,.89,1.54,6.95),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
tableSchwarzMare[[3]] <- xx2
# 1980
xx2 <- matrix(c(4.27,1.35,2.24,.7,.22,
                2.03,2.06,3.63,1.30,.27,
                3.42,3.76,21.99,9.25,4.06,
                .73,.93,5.16,9.42,7.49,
                .15,.15,1.27,2.8,11.35),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
tableSchwarzMare[[4]] <- xx2
# 1990
xx2 <- matrix(c(2.68,.68,1.31,.53,.14,
                .84,1.25,2.33,.82,.15,
                1.89,2.57,18.09,9.57,3.24,
                .68,.98,8.45,14.4,7.83,
                .14,.17,2,4.76,14.51),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
tableSchwarzMare[[5]] <- xx2
# 2000
xx2 <- matrix(c(3.47,.6,1.42,.52,.16,
                .68,1.01,1.79,.65,.13,
                1.8,2.02,15.54,7.33,2.41,
                .76,1.06,9.26,14.91,6.98,
                .17,.18,2.8,6.33,18.02),5,5,byrow=TRUE)
if(abs(sum(xx2)-100) > 0.1) {print('error')} else {print('ok')}
#xx2 <- xx2 / sum(xx2)
tableSchwarzMare[[6]] <- xx2



```


```{r}

rCSchwarzMare <- lapply(tableSchwarzMare,calcDiscreteCopula)
rHSchwarzMare <- lapply(tableSchwarzMare,calcDiscreteHellinger)
rSSchwarzMareP <- lapply(tableSchwarzMare,sort_parm)
rSSchwarzMareC <- lapply(rCSchwarzMare, function(x) sort_parm(x$pbar))
rASchwarzMare <- lapply(tableSchwarzMare,altham_index) 
#rSSchwarzMareC <- list()
#for (i in 1:length(tableSchwarzMare)){
#  rSSchwarzMareC[[i]] <- sort_parm(rCSchwarzMare[[i]]$pbar)
#}

```



```{r}

SchwarzMare_results <- matrix(0,nrow=9,ncol=6)
xyears <- c(1940,1960,1970,1980,1990,2000)
colnames(SchwarzMare_results) <- xyears
rownames(SchwarzMare_results) = c('YuleY',"HellingerAssoc",'Wtd Sort Parm Original p','Wtd Sort Parm Copula',
                                  'HellingDiagDist','Diag sum orig','Diag sum copula','Helling Depend',
                                  'Altham Ind')
#round(rCBlumeWhiteMacro$YuleY,digits=3)

xyear <- c(1940,1960,1970,1980,1990,2000)
xdiagporig <- c(0,0,0,0,0,0)
xdiagpbar <- xdiagporig

for (i in 1:6) {
  SchwarzMare_results[1,i] <- round(rCSchwarzMare[[i]]$YuleY,digits=3)
  SchwarzMare_results[2,i] <- round(rHSchwarzMare[[i]]$HellingDiagAssoc,digits=3)
  SchwarzMare_results[3,i] <- round(rSSchwarzMareP[[i]]$AvgSortParm,digits=3)
  SchwarzMare_results[4,i] <- round(rSSchwarzMareC[[i]]$AvgSortParm,digits=3)
  SchwarzMare_results[5,i] <- round(rHSchwarzMare[[i]]$HellingDiagDist,digits=3)
  SchwarzMare_results[6,i] <- round(sum(diag(rCSchwarzMare[[i]]$porig)),digits=3)
  SchwarzMare_results[7,i] <- round(sum(diag(rCSchwarzMare[[i]]$pbar)),digits=4)
  SchwarzMare_results[8,i] <- round(rHSchwarzMare[[i]]$HellingDiagAssoc,digits=3)
  SchwarzMare_results[9,i] <- round(rASchwarzMare[[i]],digits=2)
  xdiagporig[i] <- SchwarzMare_results[7,i]
  xdiagpbar[i] <- SchwarzMare_results[8,i]
}

print(SchwarzMare_results)

plot(xyear,xdiagporig)
lines(xyear,xdiagpbar)

```



## Eika, Mogstad, Zafar

In Appendix Table D$ "U.S. Marital Sorting Parameters" they give functions of the original probability matrix elements:

$$ \frac{p_{ij}}{p_{.j} p_{i.}} $$

I have to "unwind" them (in "fitSortParm.Rmd") by fitting the 6 marginals to match the matrixes of "sorting parameters"

Original "sorting parameters" (just for reference)



```{r}

vsortParm1962 <- matrix(c(1.57,.61,.40,.10,
                          .68,1.57,1.29,.84,
                          .41,.86,2.12,2.59,
                          .22,.53,1.27,4.75),4,4,byrow=TRUE)
vsortParm1980 <- matrix(c(2.59,.83,.39,.08,
                          .83,1.45,1.04,.52,
                          .32,.69,1.76,1.52,
                          .1,.28,.83,3.07),4,4,byrow=TRUE)
vsortParm2013 <- matrix(c(7.25,1,.3,.07,
                          1.13,1.96,.83,.33,
                          .52,.95,1.68,.68,
                          .11,.4,.76,1.84),4,4,byrow=TRUE)

# And then fitted probability matrixes:

plist <- list()
pEikaMogstedZafar1962 <- matrix(c(0.338337228,0.068155452,0.01959652,0.005529587,
0.129838694,0.155423079,0.05599558,0.041154478,
0.023555168,0.025616597,0.02768898,0.038180772,
0.007812158,0.009757641,0.01025229,0.043279729),4,4,byrow=TRUE)
pEikaMogstedZafar1962 <- pEikaMogstedZafar1962 / sum(pEikaMogstedZafar1962)
plist[[1]] <- pEikaMogstedZafar1962
pEikaMogstedZafar1980 <- matrix(c(0.135194922,0.05788028,0.01589550,0.004022553,
0.091928993,0.21455289,0.08994080,0.055479032,
0.014215237,0.04094918,0.06104721,0.065042743,
0.003771586,0.01410828,0.02444281,0.111535601),4,4,byrow=TRUE)
pEikaMogstedZafar1980 <- pEikaMogstedZafar1980 / sum(pEikaMogstedZafar1980)
plist[[2]] <- pEikaMogstedZafar1980
pEikaMogstedZafar2013 <- matrix(c(0.043430673,0.02092853,0.005474657,0.001856413,
0.024494472,0.14843141,0.054808117,0.031668067,
0.012036916,0.07682730,0.118467253,0.069684915,
0.003602252,0.04576374,0.075817964,0.266757828),4,4,byrow=TRUE)
pEikaMogstedZafar2013 <- pEikaMogstedZafar2013 / sum(pEikaMogstedZafar2013)
plist[[3]] <- pEikaMogstedZafar2013

rHlist <- lapply(plist,calcDiscreteHellinger)
rClist <- lapply(plist,calcDiscreteCopula)
rAlist <- lapply(plist,altham_index)



presults <- matrix(0,3,7)
rownames(presults) <- c('EikaMogstedZafar 1962','EikaMogstedZafar 1980','EikaMogstedZafar 2013')
colnames(presults) <- c("Yule's rho","Hellin Dep","Helling Assoc","Pearson",
                        "PearsonRank",'HellingDiagDist','Altham')

for (i in 1:length(plist)) {
  presults[i,1] <- round(rHlist[[i]]$YuleY,digits=4)
  presults[i,2] <- round(rHlist[[i]]$HellingDiagAssoc,digits=4)
  presults[i,3] <- round(rHlist[[i]]$HellingDiagAssoc,digits=4)
  presults[i,4] <- round(rClist[[i]]$Pearson,digits=4)
  presults[i,5] <- round(rClist[[i]]$PearsonRank,digits=4)
  presults[i,6] <- round(rHlist[[i]]$HellingDiagDist,digits=4)
  presults[i,7] <- round(rAlist[[i]],digits=3)
}
print(1)
presults
print("Sorting parameters for 2013 from discrete copula - different from original sorting parameters")
vsortParm1962 - rClist[[1]]$pbar/(.25*.25)
vsortParm2013 - rClist[[3]]$pbar/(.25*.25)
print('1962 discrete copula sorting parameters')
rClist[[1]]$pbar/(.25*.25)
print('1980 discrete copula sorting parameters')
rClist[[2]]$pbar/(.25*.25)
print('2013 discrete copula sorting parameters')
rClist[[3]]$pbar/(.25*.25)

x1 <- rowSums(pEikaMogstedZafar1962)*colSums(pEikaMogstedZafar1962)
x1 <- sum(diag(vsortParm1962)*x1) / sum(x1)
x2 <- rowSums(pEikaMogstedZafar1980)*colSums(pEikaMogstedZafar1980)
x2 <- sum(diag(vsortParm1980)*x2) / sum(x2)
x3 <- rowSums(pEikaMogstedZafar2013)*colSums(pEikaMogstedZafar2013)
x3 <- sum(diag(vsortParm2013)*x3) / sum(x3)

print(paste('Weighted diagonals of original sharing parms, 1962, 1980, 2013',x1,x2,x3))

# For discrete copula, marginals are all .25 so 
#   pbar -> sort parms requires divided by (.25*.25)
#   weighted diagonals requires  * .25
#   so overal it is daig(pbar) / .25
x1 <- sum(diag(rClist[[1]]$pbar)) / .25
x2 <- sum(diag(rClist[[2]]$pbar)) / .25
x3 <- sum(diag(rClist[[3]]$pbar)) / .25


print(paste('Weighted diagonals of discrete copula sharing parms, 1962, 1980, 2013',x1,x2,x3))


# Calculate sorting parameters using 1962 dependence and 2013 marginals
xr2013 <- rowSums(pEikaMogstedZafar2013)
xc2013 <- colSums(pEikaMogstedZafar2013)
#x1962_2013 <- ipu_matrix(rClist[[1]]$pbar, xr2013, xc2013)  # create pbar via iterative proportional fittinng, per Geenens p 430
x1962_2013 <- Ipfp(rClist[[1]]$pbar,list(1,2), list(xr2013, xc2013))$p.hat  # create pbar via iterative proportional fittinng, per Geenens p 430

print(paste("marital sorting parm for no HS degree, college, 1962 dependence and 2013 marginals",x1962_2013[1,1]/(xr2013[1]*xc2013[1]),x1962_2013[4,4]/(xr2013[4]*xc2013[4])))

# Calculate sorting parameters using 1962 marginals and 1980 and 2013 dependence
x1 <- rowSums(pEikaMogstedZafar1962)*colSums(pEikaMogstedZafar1962)
x1 <- sum(diag(vsortParm1962)*x1) / sum(x1)
xr1962 <- rowSums(pEikaMogstedZafar1962)
xc1962 <- colSums(pEikaMogstedZafar1962)
#x1980D_1962M <- ipu_matrix(rClist[[2]]$pbar, xr1962, xc1962)  # 1980 dependence, 1962 marginals
#x2013D_1962M <- ipu_matrix(rClist[[3]]$pbar, xr1962, xc1962)  # 2013 dependence, 1962 marginals
x1980D_1962M <- Ipfp(rClist[[2]]$pbar,list(1,2), list(xr1962, xc1962))$p.hat  # 1980 dependence, 1962 marginals
x2013D_1962M <- Ipfp(rClist[[3]]$pbar,list(1,2), list(xr1962, xc1962))$p.hat  # 2013 dependence, 1962 marginals
xr <- matrix(xr1962,4,4)
xc <- matrix(xc1962,4,4,byrow=TRUE)
xsort <- x1980D_1962M / (xr*xc)     # Matrix of "sorting parameters" using 1980 dependence, 1962 marginals
x2 <- xr1962*xc1962
x2 <- sum(diag(xsort)*x2) / sum(x2)
xsort <- x2013D_1962M / (xr*xc)     # Matrix of "sorting parameters" using 2013 dependence, 1962 marginals
x3 <- xr1962*xc1962
x3 <- sum(diag(xsort)*x3) / sum(x3)

print(paste("wtd avg marital sorting for 1962, 1980, 2013 using 1962 Marginals, 1962, 1980, 2013 Dependence",x1,x2,x3))

xr <- rowSums(pEikaMogstedZafar1980)
xc <- colSums(pEikaMogstedZafar1980)
#x1962D_1980M <- ipu_matrix(pEikaMogstedZafar1962,xr,xc)  # 1962 dependence, 1980 marginals
x1962D_1980M <- Ipfp(pEikaMogstedZafar1962,list(1,2),list(xr,xc))$p.hat  # 1962 dependence, 1980 marginals
x2 <- sort_parm(x1962D_1980M)$AvgSortParm
xr <- rowSums(pEikaMogstedZafar2013)
xc <- colSums(pEikaMogstedZafar2013)
#x1962D_2013M <- ipu_matrix(pEikaMogstedZafar1962,xr,xc)  # 1962 dependence, 2013 marginals
x1962D_2013M <- Ipfp(pEikaMogstedZafar1962,list(1,2),list(xr,xc))$p.hat  # 1962 dependence, 2013 marginals
x3 <- sort_parm(x1962D_2013M)$AvgSortParm
print(paste("wtd avg marital sorting for 1962, 1980, 2013 using 1962 Dependence, 1962, 1980, 2013 Marginals",x1,x2,x3))

print('Copulas for 1962, 1980, 2013')
rClist[[1]]$pbar
rClist[[2]]$pbar
rClist[[3]]$pbar

```




## Greenwood, Guner, Kocharkov, Santos 2014

Greenwood, Jeremy, Nezih Guner, Georgi Kocharkov, and Cezar Santos. 2014. “Marry Your Like: Assortative Mating and Income Inequality.” American Economic Review 104 (5): 348–53. https://doi.org/10.1257/aer.104.5.348.




```{r}


plist <- list()
pGreenwood1960 <- matrix(c(.323,.138,.019,.004,.001,
                           .076,.165,.028,.008,.002,
                           .018,.051,.027,.008,.002,
                           .005,.027,.019,.018,.003,
                           .003,.016,.017,.016,.008),5,5,byrow=TRUE)
pGreenwood1960 <- pGreenwood1960 / sum(pGreenwood1960)
plist[[1]] <- pGreenwood1960
pGreenwood2005 <- matrix(c(.039,.031,.010,.003,.001,
                           .023,.192,.082,.037,.012,
                           .005,.065,.088,.047,.016,
                           .002,.030,.045,.104,.037,
                           .001,.010,.018,.050,.053),5,5,byrow=TRUE)
pGreenwood2005 <- pGreenwood2005 / sum(pGreenwood2005)
plist[[2]] <- pGreenwood2005

rHlist <- lapply(plist,calcDiscreteHellinger)
rClist <- lapply(plist,calcDiscreteCopula)
rAlist <- lapply(plist,altham_index)


presults <- matrix(0,2,7)
rownames(presults) <- c('Greenwood 1960','Greenwood 2005')
colnames(presults) <- c("Yule's rho","Hellin Dep","Helling Assoc","Pearson","PearsonRank",'HellingDiagDist',
                        'Altham')

for (i in 1:length(plist)) {
  presults[i,1] <- round(rHlist[[i]]$YuleY,digits=4)
  presults[i,2] <- round(rHlist[[i]]$HellingDiagAssoc,digits=4)
  presults[i,3] <- round(rHlist[[i]]$HellingDiagAssoc,digits=4)
  presults[i,4] <- round(rClist[[i]]$Pearson,digits=4)
  presults[i,5] <- round(rClist[[i]]$PearsonRank,digits=4)
  presults[i,6] <- round(rHlist[[i]]$HellingDiagDist,digits=4)
  presults[i,7] <- round(rAlist[[i]],digits=3)
}  
print(1)
presults


xr <- matrix(rowSums(pGreenwood1960),5,5)
xc <- matrix(colSums(pGreenwood1960),5,5,byrow=TRUE)
psortGreenwood1960 <- pGreenwood1960 / (xr * xc)
csortGreenwood1960 <- rClist[[1]]$pbar / (.2*.2)
print(paste("1962 ratio of sum of actual to random diagonals, original and copula",sum(diag(pGreenwood1960)) / sum(diag(xr*xc)),sum(diag(rClist[[1]]$pbar)) / .2))
xr <- matrix(rowSums(pGreenwood2005),5,5)
xc <- matrix(colSums(pGreenwood2005),5,5,byrow=TRUE)
psortGreenwood2005 <- pGreenwood2005 / (xr * xc)
csortGreenwood2005 <- rClist[[2]]$pbar / (.2*.2)
print(paste("2005 ratio of sum of actual to random diagonals, original and copula",sum(diag(pGreenwood2005)) / sum(diag(xr*xc)),sum(diag(rClist[[2]]$pbar)) / .2))


x1 <- rowSums(pGreenwood1960)*colSums(pGreenwood1960)
x1 <- sum(diag(psortGreenwood1960)*x1) / sum(x1)
x2 <- rowSums(pGreenwood2005)*colSums(pGreenwood2005)
x2 <- sum(diag(psortGreenwood2005)*x2) / sum(x2)

print(paste('Weighted diagonals of original sharing parms, 1960, 2005',x1,x2))

print("1960 and 2005 sorting parameters orginal matrix")
print(psortGreenwood1960)
print(psortGreenwood2005)
print("1960 and 2005 sorting parameters copula")
print(csortGreenwood1960)
print(csortGreenwood2005)

```

